A multi-dimensional extension of the Lightweight Temporal Compression method

### 2018-11-12 paper.tex
changle long series data set and update table1 and values in Expeirment1

% packages
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\usepackage{amsfonts} % for R symbol (the set of real numbers)
\usepackage{color}
\usepackage{colortbl} % for \rowcolor
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[bookmarks=false]{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,filecolor=black,urlcolor=blue}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{subcaption}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amsthm}
\usepackage{wrapfig}
\usepackage{multicol}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

\author{Bo Li$^1$, Omid Sarbishei$^2$, Hosein Nourani$^1$, Tristan Glatard$^1$\\
  $^1$ Department of Computer Science and Software Engineering, Concordia University, Montreal, QC, Canada \\
  $^2$  Research and Development Department, Motsai Research, Saint Bruno, QC, Canada}
\maketitle

\begin{abstract}
Lightweight Temporal Compression
(LTC) is among the lossy stream
compression methods that provide the highest compression rate for the 
lowest CPU and memory consumption. As such, it is well suited to 
compress data streams in energy-constrained systems such as connected 
objects. The current formulation of LTC, however, is one-dimensional 
while data acquired in connected objects is often multi-dimensional: for instance, 
accelerometers and gyroscopes usually measure variables along 
3 directions. In this 
paper, we investigate the extension of LTC to 
higher dimensions. First, we provide a formulation 
of the algorithm in an arbitrary vectorial space of dimension $n$. 
Then, we implement the algorithm for the infinity and Euclidean norms, 
in spaces of dimension 2D+t and 3D+t. We evaluate our implementation on 
3D acceleration streams of human activities. 
Results show that the 3D implementation of LTC can save up to 20\% in 
energy consumption for low-paced activities, with a memory usage of about 100~B.
\end{abstract}

@@

With the recent technological advances in Internet of Things (IoT) 
applications, more than one billion connected objects are expected to 
be launched worldwide by 2025\footnote{\url{ 
https://www.statista.com/statistics/471264/iot-number-of-connected-devices-worldwide}}.
Power consumption is among the biggest challenges targeting connected 
objects, particularly in the industrial domains, where several sensing 
systems are commonly launched in the field to run for days or even 
weeks without being recharged. Typically, such devices use sensors to 
capture properties such as temperature or motion, and stream them to a 
host system over a radio transmission protocol such as Bluetooth 
Low-Energy (BLE). System designers aim to reduce the rate of data 
transmission as much as possible, as radio transmission is a 
power-hungry operation.

@@

Resource-intensive lossy compression algorithms such as the ones based on 
polynomial interpolation, discrete cosine and Fourier transforms, or 
auto-regression methods~\cite{lu2010optimized} are not well-suited for 
connected objects, due to the limited memory available on 
these systems (typically a few KB), and the energy consumption 
associated with CPU usage. Instead, compression algorithms need 
to find a trade-off between reducing network communications and 
increasing memory and CPU usage. As 
discussed in~\cite{zordan2014performance}, linear compression methods 
provide a very good compromise between these two factors, leading to 
substantial energy reduction.

@@

% LTC summary
% define transmitted, received points, copmression ratio
The Lightweight Temporal Compression method 
(LTC~\cite{schoellhammer2004lightweight}) has been designed 
specifically for energy-constrained systems, initially sensor networks. 
It approximates data points by a piece-wise linear function that 
guarantees an upper bound on the reconstruction error, and a reduced 
memory footprint in $\mathcal{O}(1)$. However, LTC has only been 
described for 1D streams, while streams acquired by connected objects, such as 
acceleration or gyroscopic data, are often multi-dimensional.

In this paper, we extend LTC to dimension $n$. To do so, we propose an 
algebraic formulation of the algorithm that also yields a 
norm-independent expression of it. We implement our extension on 
Motsai's Neblina module\footnote{\url{ https://motsai.com/products/neblina}}, and we test it on 3D acceleration streams 
acquired during human exercises, namely biceps curling, walking and 
running. Our implementation of LTC is available as free software.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{./figures/ltc.pdf}
\caption{Illustration of the LTC algorithm. Blue 
dots are received points, red dots are transmitted points. Dashed lines 
represent the high and low lines when a point is 
transmitted.\vspace*{-0.3cm}}
\label{fig:ltc}
\end{figure}

% Outline
Section~\ref{sec:ltc} provides some background on the LTC algorithm, 
and formalizes the description initially proposed 
in~\cite{schoellhammer2004lightweight}. Section~\ref{sec:extension} 
presents our norm-independent extension to dimension $n$, and 
Section~\ref{sec:implementation} describes our implementation. 
Section~\ref{sec:results} reports on experiments to validate our 
implementation, and evaluate the impact of n-dimensional LTC on 
energy consumption of connected objects.

@@

\begin{algorithm}
\begin{algorithmic}[1]
\Input
   \Desc{$(u^k_j, y^k_j)$}{$\quad \quad $Received data stream}
   \Desc{$\epsilon$}{$\quad \quad$Error bound}
\EndInput
\Output
   \Desc{tr}{Transmitted points}
\EndOutput
\State tr = $(u^0_0, y^0_0)$ \Comment{Last transmitted point}
\State k = 0 ; j = 1
\State (lp, hp) = ($y^0_1 - \epsilon$, $y^0_1 + \epsilon$) \Comment{Low and high points}

\While{True} \Comment{Process received points as they come}
    \State j += 1
    \State new\_lp = max($y^k_j-\epsilon$, line($u^k_j$, tr, ($u^k_{j-1}$, lp)))
    \State new\_hp = max($y^k_j+\epsilon$, line($u^k_j$, tr, ($u^k_{j-1}$, hp)))
    \If{new\_lp $\leq$ new\_hp} \Comment{Keep compressing}
        \State (lp, hp) = (new\_lp, new\_hp)
        \State \textbf{continue}
    \EndIf
    \State tr = $(u^k_{j-1}, (lp+hp)/2)$
    \Comment{Transmit point}
    \State (lp, hp) = ($y^k_j-\epsilon$, $y^k_j+\epsilon$)
    \State k += 1
    \State j = 1
\EndWhile
\end{algorithmic}
\caption{Original LTC algorithm, adapted from~\cite{schoellhammer2004lightweight}.}
\label{algo:ltc}
\end{algorithm}

@@

We note that the formulation of LTC 
in~\cite{schoellhammer2004lightweight} relies on the intersection of 
\emph{convex cones} in dimension $n+1$. For $n=1$, it corresponds to 
the intersection of triangles, which can efficiently be computed by 
maintaining boundary lines, as detailed previously. In higher 
dimension, however, cone intersections are not so straightforward, due 
to the fact that the intersection between cones may not be a cone.

@@

Let $(u_0^k, y_0^k) \in \mathbb{R}^{n+1}$ be the latest transmitted point. For convenience, all the subsequent points will be
expressed in the orthogonal space with origin $(\tau_k, \xi_k)$. We denote by $(v_j, z_j)_{j \in \llbracket 0, m \rrbracket}$ such points:
\begin{equation*}
\forall j \leq m,\  (v_j, z_j) = (u_j^k - \tau_k, y_j^k - \xi_k)
\end{equation*}
Let $\mathcal{B}_j$ be the ball of $\mathbb{R}^n$ of centre $\frac{v_1}{v_j}z_j$ and radius
$\frac{v_1}{v_j}\epsilon$:
\begin{equation*}
\mathcal{B}_j = \left\{ z \in \mathbb{R}^n,\  \norm{z-\frac{v_1}{v_j}z_j} \leq \frac{v_1}{v_j}\epsilon \right\}
\end{equation*}

@@

We define the \emph{LTC property} as follows:
\begin{equation*}
\exists z \in \mathbb{R}^n, \ \forall j \in \llbracket 1, m \rrbracket, \norm{\frac{v_j}{v_1}z-z_j} \leq \epsilon.
\end{equation*}
The original LTC algorithm ensures that the LTC property is
verified between each transmission. Indeed, all the data points
$z$ such that $(v_1, z)$ is between the high line and the low line
verify the property. Line 13 in Algorithm~\ref{algo:ltc} guarantees that
such a point exists.

@@

\State tr = ($\tau, \xi$) = ($u^0_0, y^0_0$) \Comment{Last transmitted point}
\State k = 0 ; j = 0
\While{True}
    \State j += 1
    \State ($v_j, z_j$) = ($u_j^k - \tau, y_j^k - \xi$)
    \If{$\bigcap_{l=1}^j{\mathcal{B}_l} \neq \O$} \Comment{Keep compressing}
        \State \textbf{continue}
    \EndIf
    \State Pick $z$ in $\bigcap_{l=1}^{j-1}{\mathcal{B}_j}$ \Comment{Transmit point}
    \State tr = ($\tau$, $\xi$) = ($u^k_{j-1}, z$)
    \State k += 1
    \State j = 1
\EndWhile
\end{algorithmic}
\caption{Generalized LTC.}
\label{algo:general-ltc}
\end{algorithm}

@@

Although Algorithm~\ref{algo:general-ltc} looks simple, one should not
overlook the fact that there is no good general algorithm to test
whether a set of balls intersect. The best general algorithm we could find
so far relies on Helly's theorem which is formulated as follows~\cite{helly1923mengen}:
\begin{theorem}
Let $\left\{ X_i \right\}_{i \in \llbracket 1, m \rrbracket}$ be a collection of convex subsets of $\mathbb{R}^n$. If the intersection of every $n+1$
subsets is non-empty, then the whole collection has an non-empty intersection.
\end{theorem}
\noindent This theorem leads to an algorithm of complexity ${m \choose n+1}$ which is
not usable in resource-constrained environments.

@@

As mentioned before, norm selection in $\mathbb{R}^n$ has a critical
impact on the compression error and ratio. To appreciate this effect,
let us compare the infinity norm and the
Euclidean norm in dimension 2. By comparing the unit disk to a
square of side 2, we obtain that the compression ratio of a random stream would
be $\frac{4}{\pi}$ times larger with the infinity norm than with the 
Euclidean norm. In 3D, this ratio would be $\frac{6}{\pi}$. Conversely, 
a compression error bounded by $\epsilon$ with the infinity norm 
corresponds to a compression error of $\frac{\epsilon}{\sqrt{n}}$ with 
the Euclidean norm. Unsurprisingly, the
infinity norm is more tolerant than the Euclidean norm.

@@

To implement LTC in nD with the infinity norm, we maintain a cuboid 
representation of $\cap_{l=1}^j{\mathcal{B}_l}$ throughout the 
iterations of the \texttt{while} loop in 
Algorithm~\ref{algo:general-ltc}. The implementation works with
constant memory and requires limited CPU time.

@@

\If{$\exists\ \mathcal{B}_i \in S$ s.t. $\mathcal{B}_j \cap \mathcal{B}_i = \O$}
\State \Return ($S$, False) \Comment{Some balls don't intersect}
\EndIf
\If{$\mathcal{B}_j \cap B = \O$} \Comment{Ball is outside bounding box}
\State \Return ($S$, $B$, False) 
\EndIf
\If{$\exists\ \mathcal{B}_i \in S$ s.t. $\mathcal{B}_j \subset \mathcal{B}_i$} \Comment{Remove inclusions}
\State Remove $\mathcal{B}_i$ from $B$. Add $\mathcal{B}_j$ to $B$.
\EndIf
\State $B$ = box($\mathcal{B}_j$) $\ \bigcap\ $ $B$
\State $S$ = $S \ \bigcup \  \{\mathcal{B}_j\}$
\State $x$ = find\_bisection($S$, $B$) \Comment{This can take some time}
\If{$x$ == Null}
\State \Return ($S$, $B$, False)
\Else
\State \Return ($S$, $B$, True)
\EndIf
\end{algorithmic}
\caption{Intersection test for Euclidean n-balls.}
\label{algo:euclidean}
\end{algorithm}

@@

\begin{wrapfigure}{R}{3cm}
\includegraphics[width=3cm]{figures/neblina-wrist.png}
\caption{Setup with Neblina.}
\label{fig:neblina-wrist}
\end{wrapfigure}
We conducted two experiments using Motsai's Neblina 
module, a system 
with a Nordic Semiconductor nRF52832 micro-controller, 64~KB of RAM, 
and Bluetooth Low Energy connectivity. Neblina has a 3D 
accelerometer, a 3D gyroscope, a 3D magnetometer, and environmental 
sensors for humidity, temperature and pressure. The platform is 
equipped with sensor fusion algorithms for 3D orientation tracking and 
a machine learning engine for complex motion analysis and motion 
pattern recognition~\cite{sarbishei2016accuracy}. Neblina has a 
battery of 100mAh; at 200~Hz, its average consumption is 2.52~mA when using 
accelerometer and gyroscope sensors but without radio 
transmission, and 3.47~mA with radio transmission, leading to an 
autonomy of 39.7~h without transmission and 28.8~h with transmission.

@@

We validated the behaviour of our LTC extension on a PC using data 
acquired with Neblina. We collected two 3D accelerometer time-series, a 
short one and a longer one, acquired on two different subjects 
performing biceps curl, with a 50~Hz sampling rate (see 
Figure~\ref{fig:datasets-1}). In both cases, the subject was wearing 
Neblina on their wrist, as in Figure~\ref{fig:neblina-wrist}. It should be noted that the longer 
time-series also has a higher amplitude, perhaps due to differences between 
subjects.

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-x.pdf}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-y.pdf}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-z.pdf}
\caption{Short biceps curl$^a$}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-x.pdf}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-y.pdf}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-z.pdf}
\caption{Long biceps curl}
\end{subfigure}
\caption{Time-series used in Experiment 1\\
{\footnotesize $^a$ Average of az data is -7.81g. It was shifted to 0 so that the graphs can all use the same y scale.}}
\label{fig:datasets-1}
\end{figure*}

We compressed the time-series with various values of $\epsilon$, using 
our 2D (x and y) and 3D (x, y and z) implementations of LTC. On 
Neblina, the raw uncalibrated accelerometer data corresponds to errors 
around 20~mg (1~g is 9.8~m/s$^2$). We used a 
laptop computer with 16~GB of RAM, an Intel i5-3210M CPU @ 2.50GHz 
$\times$ 4, and Linux Fedora 27. We measured memory consumption using 
Valgrind's massif 
tool~\cite{nethercote2006building}\footnote{\url{http://valgrind.org}}, 
and processing time using \texttt{gettimeofday()} from the GNU C 
Library.

Results are reported in Table~\ref{table:results-validation}. 
As expected, the compression ratio increases with $\epsilon$, and the 
maximum measured error remains lower than $\epsilon$ in all cases. The 
maximum is reached most of the time on these time-series.

\paragraph{Infinity vs Euclidean norms}
The average ratio between the compression ratios obtained
with the infinity and Euclidean norm is 1.08 for 2D data, and 1.20
for 3D data. These ratios are lower than the theoretical values of
$\frac{4}{\pi}$ in 2D and $\frac{6}{\pi}$ in 3D, which are obtained for
random-uniform signals. Unsurprisingly, the infinity norm surpasses the
Euclidean norm in terms of resource consumption. Memory-wise, the
infinity norm requires a constant amount of 80~B, used to store the
intersection of n-balls. The Euclidean norm, however, uses up to 4.9~KB of memory
for the Long time-series in 3D with $\epsilon$=48.8~mg. More importantly,
the amount of required memory increases for more complex time-series
(Long vs Short), and it also increases with larger
values of $\epsilon$. Similar observations are made for the processing
time, with values ranging from 0.2~ms for the simplest time-series and
smallest $\epsilon$, to 26.4~ms for the most complex time-series and
largest~$\epsilon$.
%~ Figure~\ref{fig:memory} shows the memory consumption 
%~ of the 3D Euclidean implementation for $\epsilon$=48.8~mg: peaks appear
%~ at the end of long compressed ranges where the signal was $\epsilon$-closed
%~ to the linear approximation.

\paragraph{2D vs 3D}
For a given $\epsilon$, the compression 
ratios are always higher in 2D than in 3D. It makes sense since the
probability for the signal to deviate from a straight line
approximation is higher in 3D than it is in 2D. Besides, resource
consumption is higher in 3D than in 2D: for the infinity norm, 3D
consumes 1.4 times more memory than 2D (1.6 times on average for
Euclidean norm), and processing time is 1.3 longer (1.18 on
average for Euclidean norm).

>>>>

% packages
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\usepackage{amsfonts} % for R symbol (the set of real numbers)
\usepackage{color}
\usepackage{colortbl} % for \rowcolor
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[bookmarks=false]{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,filecolor=black,urlcolor=blue}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{subcaption}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amsthm}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

\author{Bo Li$^1$, Omid Sarbishei$^2$, Tristan Glatard$^1$\\
  $^1$ Department of Computer Science and Software Engineering, Concordia University, Montreal, QC, Canada \\
  $^2$  Research and Development Department, Motsai Research, Saint Bruno, QC, Canada}
\maketitle

\begin{abstract}
Lightweight Temporal Compression
(LTC) is among the lossy stream
compression methods that provide the highest compression rate for the 
lowest CPU and memory consumption. As such, it is very well suited to 
compress data streams in energy-constrained systems such as connected 
objects. The current formulation of LTC, however, is one-dimensional 
while data acquired in connected objects is often multi-dimensional: for instance, 
accelerometers and gyroscopes usually measure variables along 
3 directions related by the actual movement of the sensor. In this 
paper, we investigate the extension of LTC to 
higher dimensions. First, we provide a formulation 
of the algorithm in an arbitrary vectorial space of dimension $n$. 
Then, we implement the algorithm for the infinity and Euclidean norms, 
in spaces of dimension 2D+t and 3D+t. We evaluate our implementation on 
3D acceleration streams acquired during human activities. 
Results show that the 3D implementation of LTC can save up to 20\% in 
energy consumption for low-paced activities, with a memory usage of about 100~B.
\end{abstract}

@@

With the recent technological advances in Internet of Things (IoT) 
applications, more than one billion connected objects are expected to 
be launched worldwide by 2025\footnote{\url{ 
https://www.statista.com/statistics/471264/iot-number-of-connected-devices-worldwide}} 
Power consumption is amongst the biggest challenges targeting connected 
objects, particularly in the industrial domains, where several sensing 
systems are commonly launched in the field to run for days or even 
weeks without being recharged. Typically, such devices use sensors to 
capture properties such as temperature or motion, and stream them to a 
host system over a radio transmission protocol such as Bluetooth 
Low-Energy (BLE). System designers aim to reduce the rate of data 
transmission as much as possible, as radio transmission is a 
power-hungry operation.

@@

Resource-intensive lossy compression algorithms such as the ones based on 
polynomial interpolation, discrete cosine and Fourier transforms, or 
auto-regression methods~\cite{lu2010optimized} are not well-suited for 
energy-constrained systems, due to the limited memory available on 
these systems (typically a few KB), and the energy consumption 
associated with CPU usage. On such systems, compression algorithms need 
to find a trade-off between reducing network communications and 
increasing memory and CPU usage. As 
discussed in~\cite{zordan2014performance}, linear compression methods 
provide a very good compromise between these two factors, leading to 
substantial energy reduction.

@@

% LTC summary
% define transmitted, received points, copmression ratio
The Lightweight Temporal Compression method 
(LTC~\cite{schoellhammer2004lightweight}) has been designed 
specifically for energy-constrained systems, initially sensor networks. 
It approximates data points by a piecewise linear function that 
guarantees an upper bound on the reconstruction error, and a reduced 
memory footprint in $\mathcal{O}(1)$. However, LTC has only been 
described for 1D streams, while streams acquired by connected objects, such as 
acceleration or gyroscopic data, are often multi-dimensional.

In this paper, we extend LTC to dimension $n$. To do so, we propose an 
algebraic formulation of the algorithm that also yields a 
norm-independent expression of it. We implement our extension on 
Motsai's Neblina module, and we test it on 3D acceleration streams 
acquired during human exercises, namely biceps curling, walking and 
running. Our implementation of LTC is available as free software.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{ltc.pdf}
\caption{Illustration of the LTC algorithm. Blue 
dots are received points, red dots are transmitted points. Dashed lines 
represent the high and low lines when a point is 
transmitted.\vspace*{-0.3cm}}
\label{fig:ltc}
\end{figure}

\todo{show $\xi$ in the Figure}

% Outline
Section~\ref{sec:ltc} provides some background on the LTC algorithm, 
and formalizes the description initially proposed 
in~\cite{schoellhammer2004lightweight}. Section~\ref{sec:extension} 
presents our norm-independent extension to dimension $n$, and 
Section~\ref{sec:implementation} describes our implementation. 
Section~\ref{sec:results} reports on experiments to validate our 
implementation, and evaluates the impact of n-dimensional LTC on 
energy consumption of connected objects.

@@

\begin{algorithm}
\begin{algorithmic}[1]
\Input
   \Desc{$(u^k_j, y^k_j)$}{$\quad \quad $Received data stream}
   \Desc{$\epsilon$}{$\quad \quad$Error bound}
\EndInput
\Output
   \Desc{tr}{Transmitted points}
\EndOutput
\State tr = $(u^0_0, y^0_0)$ \Comment{last transmitted point}
\State k = 0 ; j = 1
\State (lp, hp) = ($y^0_1 - \epsilon$, $y^0_1 + \epsilon$) \Comment{low and high points}

\While{True} \Comment{Process received points as they come}
    \State j += 1
    \State new\_lp = max($y^k_j-\epsilon$, line($u^k_j$, tr, ($u^k_{j-1}$, lp)))
    \State new\_hp = max($y^k_j+\epsilon$, line($u^k_j$, tr, ($u^k_{j-1}$, hp)))
    \If{new\_lp $\leq$ new\_hp} \Comment{Keep compressing}
        \State (lp, hp) = (new\_lp, new\_hp)
        \State \textbf{continue}
    \EndIf
    \State tr = $(u^k_j, (lp+hp)/2)$
    \Comment{Transmit point}
    \State (lp, hp) = ($y^k_j-\epsilon$, $y^k_j+\epsilon$)
    \State k += 1
    \State j = 1
\EndWhile
\end{algorithmic}
\caption{Original LTC algorithm, adapted from~\cite{schoellhammer2004lightweight}.}
\label{algo:ltc}
\end{algorithm}

@@

We note that the formulation of LTC in~\cite{schoellhammer2004lightweight} relies on
the intersection of \emph{convex cones} in dimension $n+1$. For $n=1$, it
corresponds to the intersection of triangles, which can efficiently be
computed by maintaining boundary lines, as detailed previously. In higher dimension, however, cone intersections are not so
straighforward, due to the fact that the intersection between cones
may not be a cone.

@@

Let $(u_0^k, y_0^k) \in \mathbb{R}^{n+1}$ be the latest transmitted point. For convenience, all the subsequent points will be
expressed in the orthogonal space with origin $(\tau_k, \xi_k)$. We denote by $(v_j, z_j)_{j \in \llbracket 0, m \rrbracket}$ such points:
\begin{equation*}
\forall j \leq m,\  (v_j, z_j) = (u_j^k - \tau_k, y_j^k - \xi_k)
\end{equation*}
Let $\mathcal{B}_j$ be the ball of $\mathbb{R}^n$ of center $\frac{v_1}{v_j}z_j$ and radius
$\frac{v_1}{v_j}\epsilon$:
\begin{equation*}
\mathcal{B}_j = \left\{ z \in \mathbb{R}^n,\  \norm{z-\frac{v_1}{v_j}z_j} \leq \frac{v_1}{v_j}\epsilon \right\}
\end{equation*}

@@

We define the \emph{LTC property} as follows:
\begin{equation*}
\exists z \in \mathbb{R}^n, \ \forall j \in \llbracket 1, m \rrbracket, \norm{\frac{v_j}{v_1}z-z_j} \leq \epsilon.
\end{equation*}
The original LTC algorithm ensures that the LTC property is
verified between each transmission. Indeed, all the data points
$z$ such that $(v_1, z)$ is between the high line and the low line
verify the property. Line 9 in Algorithm~\ref{algo:ltc} guarantees that
such a point exists.

@@

\State tr = ($\tau, \xi$) = ($u^0_0, y^0_0$)
\State k = 0 ; j = 0
\While{True}
    \State j += 1
    \State ($v_j, z_j$) = ($u_j^k - \tau, y_j^k - \xi$)
    \If{$\bigcap_{l=1}^j{\mathcal{B}_l} \neq \O$}
        \State \textbf{continue}
    \EndIf
    \State Pick $z$ in $\bigcap_{l=1}^{j-1}{\mathcal{B}_j}$
    \State tr = ($\tau$, $\xi$) = ($u^k_{j-1}, z$)
    \State k += 1
    \State j = 0
\EndWhile
\end{algorithmic}
\caption{Generalized LTC}
\label{algo:general-ltc}
\end{algorithm}

\todo{Check algorithms (reimplement)}

@@

Although Algorithm~\ref{algo:general-ltc} looks simple, one should not
overlook the fact that there is no good general algorithm to test
whether a set of balls intersect. The best general algorithm we could find
so far relies on Helly's theorem which is formulated as follows~\cite{helly1923mengen}:
\begin{theorem}
Let $\left\{ X_i \right\}_{i \in \llbracket 1, m \rrbracket}$ be a collection of convex subsets of $\mathbb{R}^n$. If the intersection of every $n+1$
subsets is non-empty, then the whole collection has an non-empty intersection.
\end{theorem}
\noindent This theorem leads to an algorithm of complexity ${m \choose n+1}$ which is
not usable in resource-constrained environment.

@@

As mentioned before, norm selection in $\mathbb{R}^n$ has a critical
impact on the compression error and ratio. To appreciate this effect,
let us compare the infinity norm and the
Euclidean norm in dimension 2. By comparing the unit disk to a
square of side 2, we obtain that the compression ratio of a random stream would
be $\frac{4}{\pi}$ times larger with the infinity norm than with the 
Euclidean norm. In 3D, this ratio would be $\frac{6}{\pi}$. Conversely, 
a compression error bounded by $\epsilon$ with the infinity norm 
corresponds to a compression error of $\frac{\epsilon}{\sqrt{n}}$ with 
the Euclidean norm. Unsurprisingly, the
infinity norm is more `tolerant' than the Euclidean norm.

@@

To implement LTC with the infinity norm, we maintain a cuboid 
representation of $\cap_{l=1}^j{\mathcal{B}_l}$ throughout the 
iterations of the \texttt{while} loop in 
Algorithm~\ref{algo:general-ltc}. The implementation works with
constant memory and requires limited CPU time.

@@

\If{$\exists\ \mathcal{B}_i \in S$ s.t. $\mathcal{B}_j \cap \mathcal{B}_i = \O$}
\State \Return ($S$, False)
\EndIf
\If{$\mathcal{B}_j \cap B = \O$}
\State \Return ($S$, $B$, False)
\EndIf
\If{$\exists\ \mathcal{B}_i \in S$ s.t. $\mathcal{B}_j \subset \mathcal{B}_i$}
\State Remove $\mathcal{B}_i$ from $B$. Add $\mathcal{B}_j$ to $B$.
\EndIf
\State $B$ = box($\mathcal{B}_j$) $\ \bigcap\ $ B 
\State $S$ = $S \ \bigcup \  \{\mathcal{B}_j\}$
\State $x$ = find\_bisection($S$, $B$)
\If{$x$ == Null}
\State \Return ($S$, $B$, False)
\Else
\State \Return ($S$, $B$, True)
\EndIf
\end{algorithmic}
\caption{Intersection test for Euclidean n-balls.}
\label{algo:euclidean}
\end{algorithm}

@@

We conducted two experiments using Motsai's Neblina 
module\footnote{\url{ https://motsai.com/products/neblina}}, a system 
with a Nordic Semiconductor nRF52832 micro-controller, 64~KB of RAM, 
and Bluetooth Low Energy connectivity. Neblina has a 3D 
accelerometer, a 3D gyroscope, a 3D magnetometer, and environmental 
sensors for humidity, temperature and pressure. The platform is 
equipped with sensor fusion algorithms for 3D orientation tracking and 
a machine learning engine for complex motion analysis and motion 
pattern recognition~\cite{sarbishei2016accuracy}. Neblina has a 
battery of 100mAh; its average consumption is 2.52~mA when using 
accelerometer and gyroscope sensors at 200~Hz but without radio 
transmission, and 3.47~mA with radio transmission, leading to an 
autonomy of 39.7~h without transmission and 28.8~h with transmission.

@@

We validated the behavior of our LTC extension on a 
PC using data acquired with Neblina. We collected two 3D 
accelerometer time series, a short one and a longer one, acquired 
on two different subjects performing biceps curl, with a 50Hz sampling rate (see 
Figure~\ref{fig:datasets-1}). In both cases, the 
subject was wearing Neblina on their right wrist. It should be noted that the longer dataset also has 
a higher amplitude, most likely due to differences between subjects.

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-x.pdf}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-y.pdf}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-z.pdf}
\caption{Short biceps curl$^a$}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-x.pdf}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-y.pdf}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-z.pdf}
\caption{Long biceps curl}
\end{subfigure}
\caption{Time series used in Experiment 1\\
{\footnotesize $^a$ Average of az data is -7.81g. It was shifted to 0 so that the graphs can all use the same y scale.}}
\label{fig:datasets-1}
\end{figure*}

We compressed the time series with various values of $\epsilon$, using 
our 2D (x and y) and 3D (x, y and z) implementations of LTC. On 
Neblina, the raw uncalibrated accelerometer data corresponds to errors 
around 20~mg (1~g is 9.8~m/s$^2$). We used a 
laptop computer with 16~GB of RAM, an Intel i5-3210M CPU @ 2.50GHz 
× 4, and Linux Fedora 27. We measured memory consumption using 
Valgrind's massif 
tool~\cite{nethercote2006building}\footnote{\url{http://valgrind.org}}, 
and processing time using \texttt{gettimeofday()} from the GNU C 
Library.

Results are reported in Table~\ref{table:results-validation}. 
As expected, the compression ratio increases with $\epsilon$, and the 
maximum measured error remains lower than $\epsilon$ in all cases. The 
maximum is reached most of the time on these time series.

\paragraph{Infinity vs Euclidean norms}
Compared to the infinity norm, the average ratio between the compression ratios obtained
with the infinity and Euclidean norm is 1.08 for 2D data, and 1.20
for 3D data. These ratios are lower than the theoretical values of
$\frac{4}{\pi}$ in 2D and $\frac{6}{\pi}$ in 3D, which are obtained for
random-uniform signals. Unsurprisingly, the infinity norm surpasses the
Euclidean norm in terms of resource consumption. Memory-wise, the
infinity norm requires a constant amount of 80~B, used to store the
intersection of n-balls. The Euclidean norm, however, uses up to 4.7~KB of memory
for the Long time series in 3D with $\epsilon$=48.8~mg. More importantly,
the amount of required memory increases for more complex time series
(Long vs Short), and it also increases with larger
values of $\epsilon$. Similar observations are made for the processing
time, with values ranging from 0.2~ms for the simplest time series and
smallest $\epsilon$, to 41.3~ms for the most complex time series and
largest~$\epsilon$.
%~ Figure~\ref{fig:memory} shows the memory consumption
%~ of the 3D Euclidean implementation for $\epsilon$=48.8~mg: peaks appear
%~ at the end of long compressed ranges where the signal was $\epsilon$-closed
%~ to the linear approximation.

\paragraph{2D vs 3D}
For a given $\epsilon$, the compression
ratios are always higher in 2D than in 3D. It makes sense since the
probability for the signal to deviate from a straight line
approximation is higher in 3D than it is in 2D. Besides, resource
consumption is higher in 3D than in 2D: for the infinity norm, 3D
consumes 1.4 times more memory than 2D (1.3 times on average for
Euclidean norm), and processing time is 1.2 longer (1.13 on
average for Euclidean norm).
